Scalar and Representative Observables for Polarimetric SAR Data

### 2013-09-30 paper_dev/article.tex
abc d

In this section, the generic scalar statistical model for POLSAR is presented.
After the basic foundations of POLSAR statistical analysis is introduced,
  this section presents the theoretical model for multi-dimensional POLSAR data.

In this paper, the POLSAR scattering vector is denoted as $s$.
In the case of partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
In the case of full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry,
and $L$ is the number of looks.

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behave like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

We will now use this result to develop the generic scalar statistical model for POLSAR. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} %and \ref{eqn:sum_log_chi_squared_rv}
we have: %can deduce the following relationships:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this article, the determinant-ratio is also proposed as a discrimination measure for the POLSAR data,
  which also have just been shown to suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.
Assuming, on the one hand,
  that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
then the determinant-ratio of the signal random variable ($\mathbb{R}_{\Sigma}$) %and log-distance ($\mathbb{L}$)
%  are observable according to the following definitions:
  is defined as:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R}_{\Sigma} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}

On the other hand, under a looser assumption %From another perspective
  where the POLSAR data is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
 then random variable called the change-ratio ($\mathbb{R}_{C}$)  is defined as:
\begin{eqnarray}
  \mathbb{R}_{C} &=& \frac{|C_1|}{|C_2|} \label{eqn:determinant_ratio_samples_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
where $C_1$ and $C_2$ are samples of the covariance matrix determinant in an assumed homogeneous area.

@@

Using the results from Eqn. \ref{eqn:determinant_distribution}, %\ref{eqn:log_determinant_distribution} and \ref{eqn:avg_log_det}
  we have
\begin{eqnarray}
\mathbb{R}_{\Sigma} &\sim& \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_ratio_distribution} \\
\mathbb{R}_{C} &\sim& \prod_{i=0}^{d-1} \frac{\chi(2L-2i)}{\chi(2L-2i)}
%\mathbb{L} &\sim&  \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln(2L)
%\label{eqn:log_determinant_distance_distribution} \\ 
% \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2} + k
%\label{eqn:dispersion_distribution} \\ 
% \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
%\label{eqn:contrast_distribution}  
\end{eqnarray}
%with $\Delta(2L) \sim \Lambda(2L) - \Lambda(2L)$
%and $k=\sum^{d-1}_{i=0} \psi^0(L-i)$

@@

Since each elementary component follows fixed distributions (i.e. $\chi^2(2L)$),
  it is natural that this variable also follows fixed distributions.
Moreover, it is independent of the underlying signal $\Sigma_v$,
  indicating its statistically consistent properties and hence its applicability as a POLSAR discrimination measure.
  This claim is extrapolated from the widely-used intensity-ratio as SAR discrimination measure and is further strengthened in the next section which shows that the generic models for POLSAR presented in this paper also include the commonly-used models for SAR as its special 1-dimensional case. 
%This result shows how
%In short, these random variables are shown to follow consistent and fixed distributions,
%  regardless of the underlying signal $\Sigma_v$.

@@

The previous sections have introduced and validated the theoretical models for 3-dimensional ($d=3$) full polarimetric and 2-dimensional ($d=2$) partial polarimetry, cases.
This section will show that the model is equally  applicable to the 1-dimensional ($d=1$) case,
  which is physically equivalent to  collapsing the multi-dimensional POLSAR dataset  into conventional single dimensional SAR data.
Mathematically, the sample covariance matrix is reduced to the sample variance while the determinant is equivalent to  the scalar value.
As variance is equal to intensity for SAR data, our result is consistent with previous results for SAR intensity data.
Hence this derivation can be considered
  either as cross-validation for the proposed POLSAR model
  or as evidence showing that
    the generic POLSAR model proposed in this article also covers the traditional model for SAR intensity in its natural coverage.
%  as well as reminding us that SAR is a special case of POLSAR.

@@

Upon setting $d=1$ into the above equations,
  the becomes:
\begin{align*}
  |C_v| &\sim  \frac{|\Sigma_v|}{(2L)} \cdot \chi (2L) \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim \frac{1}{(2L)} \cdot \chi(2L) \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
\mathbb{R}_{C} = \frac{C_1}{C_2} &\sim \prod_{i=0}^{d-1} \frac{\chi(2L-2i)}{\chi(2L-2i)}
\end{align*}

@@

\section{Discussion}

\subsection{Validating the proposed models against real-life data}
\label{sec:polsar_models_validation}

This section aims to verify the theoretical models given in Eqns. \ref{eqn:determinant_distribution} and \ref {eqn:determinant_ratio_distribution} against practical data.
Both of these models require the estimation of two parameters from some given captured data.
They are the dimensional number $d$ and the look number $L$.
The dimensional number $d$ is related to the type of (POL)SAR data,
  with $d=1,2,3$ corresponding to the cases of SAR, partial and full polarimetric SAR data, respectively.
The look-number $L$ can be taken either as nominally stated by the data provider,
  or it can be estimated using the technique proposed by \cite{Anfinsen_2009_TGRS_3795}.
To show the robustness of the proposed models, their validation is also put to test on different types of POLSAR sensors.
Specifically in this article, the air-borne four-look ($L=4$) AIRSAR image of Flevoland and
  a fine-quad single-look complex RADARSAT2 image ($L=1$) are employed.
Since the determinant of the covariance matrix is only significant on multi-look data,
  nine-look processing is first applied on the single-look RADARSAT2 data ($L=9$).

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  which have been validated extensively, the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_multi_look_SAR_dispersion_contrast_models} presents the results of an experiment carried out for the stated purpose.
In the experiment, the intensity of single-channel SAR data (HH) for a homogeneous area in the AIRSAR Flevoland dataset is extracted.
The histograms for both the intensity and intensity ratio (TODO:Replot) %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
The plots are obtained with ENL set to both the nominal number of 4 and the estimated number of (TODO:CALC).
Also presented are the plots of the same experiment applied on a homogeneous patch inside the nine-look processed RADARSAT2 image (TODO:NewPlot).
In this case, even though the RADARSAT2 has a nominal look-number of 9, its ``true'' value is estimated as $L=$ (TODO:CALC).
In all cases, good visual match is apparent, which validates the model proposed above. %providing us a simple validation of the result.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[AIRSAR (HH) determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_determinant_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_determinant}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_determinant_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_determinant}
	} \\
	\subfloat[AIRSAR (HH) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_det_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_det_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	} \\
	\subfloat[AIRSAR (HH) change ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_change_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) change ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	}
\end{tabular}
\caption{The specific d=1 models are validated on both practical RADARSAT2 and AIRSAR datasets}
\label{fig:verify_det_ratio_model_1D}
\end{figure}

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar experiments like the one described for SAR data is carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image using the technique described in \cite{Anfinsen_2009_TGRS_3795} and is also noted in the plots.

@@

\subsection{Theoretical Implications of This Proposal}
\label{sec:discussion}

Let us begin by noting a few theoretical properties of the proposed statistical models.
First, the use of covariance matrix log-determinant may be related to the standard eigen-decomposition method of the POLSAR covariance matrices.
In fact, the log-determinant can also be computed as the sum of log-eigenvalues.
Specifically $\ln{|M|} = \sum \ln{\lambda_M}$ where $\lambda_M$ denotes all the eigenvalues of M.
Thus similar to other eigenvalue based approach (e.g. entropy/anisotropy, ...),
  the models presented here are invariant to polarization basis transformations.

@@

Intuitively, the concept of the determinant being representative of the POLSAR data can be thought of as being similar to the concept of using the magnitude as being representative of a complex number
(in a sense this is also why intensity is widely considered representative of the SAR data complex).
Similar to the determinant, the magnitude observable is also scalar and is not lossless.
To fully represent the complex number, another observable is needed besides the magnitude.
The magnitude however is much more widely used, as it can be shown that its is invariant to a change in the reference frame.
Similarly, the determinant of the POLSAR covariance matrix is also scalar and invariant to a change of polarization basis.
Thus while it is not fully representative of the POLSAR data, it can be said to be highly representative of it.
The important thing is that, it can be used to give a scalar comparison of the multi-dimensional data.
This is also similar to the standard use of the magnitude when two or more complex numbers are to be compared. %, most of the times the magnitude is to be used.

This leads to the application of the models in deriving discrimination measures.
Besides the above properties, the theoretical models may also provide an alternative derivation for the widely used likelihood test statistics in POLSAR.
In view of the models given in Eqns \ref{eqn:determinant_distribution} \& \ref{eqn:log_determinant_distribution},
  the likelihood test statistics exposed in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be simulated as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
As a by-product of this exact derivation, the current paper also proposed several simpler discrimination measures for the common case of $L_x=L_y$.

Similar to the way that other measures of distance can be used to derive POLSAR classifiers \cite{Lee_1999_TGRS}, change detectors \cite{Conradsen_2003_TGRS_4}, edge detectors \cite{Schou_2003_TGRS_20} or other clustering and speckle filtering techniques \cite{Le_2010_ACRS} \cite{Le_2011_ACRS}, 
new detection, classification, clustering or speckle filtering algorithms can be derived using the models presented in this paper.
These applications however are not included in this paper, but rather left as a suggestion for future investigations.
There are a few reasons for this decision.
The first is that the advantages of the proposed models have been demonstrated through a few applications in this paper:
  the first being a faster ENL evaluation for POLSAR data and
  the second being the visual evaluation of POLSAR speckle filters.
The second reason is that such an application would be quite big in scope, and as such would easily distract the reader from the main proposal of this paper.
Last but not least, this paper focuses on the proposal of several scalar and representative statistical models for POLSAR, which results in associated statistical models and discrimination measures.
It does not, technically speaking, focuses on proposing new discrimination measures.

@@

%\appendices
%\section{Homoskedastic Model for the Log-Determinant}
%\label{chap:appendix_a}
%
%\subsection{Log-Chi-Square Distribution and its Derivatives}
%%\section{Log-Chi-Square Distribution}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\setcounter{equation}{0}
%
%This section provides the mathematical derivations for the log-transformed version of chi-squared random variables.
%
%Chi-squared random variables $\chi\ \sim\ \chi^2(k)\ $ follows the pdf:
%\begin{equation}
%pdf(\chi;\,k) =
%  \frac{\chi^{(k/2)-1} e^{-\chi/2}}{2^{k/2} \Gamma\left(\frac{k}{2}\right)}  
%\label{eqn:chi_squared_dist_pdf:appdixA}
%\end{equation}
%
%Setting L=k/2 into Eqn. \ref{eqn:chi_squared_dist_pdf:appdixA}
%\begin{equation}
%pdf(\chi) = \frac{\chi^{L-1}e^{-\chi/2}}{2^L\Gamma(L)}
%\end{equation}
%
%Applying the variable change theorem, which states that: if $y=\phi(x)$ with $\phi(c)=a$ and $\phi(d)=b$, then:
%\begin{equation}
% \int_a^b \! f(y) \, dy = \int_c^d \! f[\phi(x)] \frac{d\phi}{dx} dx
%\end{equation}
%into the log-transformation, which changes the random variables $\Lambda=ln(\chi)$, we have:
%\begin{eqnarray*}
%  d\chi &=& e^\Lambda d\Lambda \\

>>>>

%In this section, the generic scalar statistical model for POLSAR is presented.
In this section, after the basic foundations of POLSAR statistical analysis is introduced,
  the generic scalar statistical model for the multi-dimensional POLSAR is presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data.

In this paper, the POLSAR scattering vector is denoted as $s$.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
  and $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry.

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
This result is used here to develop the generic scalar statistical model for POLSAR. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} %and \ref{eqn:sum_log_chi_squared_rv}
we have: %can deduce the following relationships:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data.
  which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

%Assuming, on the one hand,
For cases where the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
then the determinant-ratio of the signal random variable ($\mathbb{R}_{\Sigma}$) %and log-distance ($\mathbb{L}$)
%  are observable according to the following definitions:
  is defined as:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R}_{\Sigma} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
%On the other hand, under a looser assumption %From another perspective
For cases where the POLSAR data is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
 then random variable called the change-ratio ($\mathbb{R}_{C}$)  is defined as:
\begin{eqnarray}
  \mathbb{R}_{C} &=& \frac{|C_1|}{|C_2|} \label{eqn:determinant_ratio_samples_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
where $C_1$ and $C_2$ are samples of the covariance matrix determinant in an assumed homogeneous area.

@@

Using the results from Eqn. \ref{eqn:determinant_distribution}, %\ref{eqn:log_determinant_distribution} and \ref{eqn:avg_log_det}
  we have
\begin{eqnarray}
\mathbb{R}_{\Sigma} &\sim& \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_ratio_distribution} \\
\mathbb{R}_{C} &\sim& \prod_{i=0}^{d-1} \frac{\chi(2L-2i)}{\chi(2L-2i)} \label{eqn:change_ratio_distribution}
%\mathbb{L} &\sim&  \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln(2L)
%\label{eqn:log_determinant_distance_distribution} \\ 
% \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2} + k
%\label{eqn:dispersion_distribution} \\ 
% \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
%\label{eqn:contrast_distribution}  
\end{eqnarray}
%with $\Delta(2L) \sim \Lambda(2L) - \Lambda(2L)$
%and $k=\sum^{d-1}_{i=0} \psi^0(L-i)$

@@

Since each elementary component follows fixed distributions (i.e. $\chi^2(2L)$),
  it is natural that this variable also follows fixed distributions.
Moreover, it is independent of the underlying signal $\Sigma_v$,
  indicating its statistically consistent properties and hence its applicability as a POLSAR discrimination measure.
This claim is extrapolated from the widely-used intensity-ratio as SAR discrimination measure
  and will be further discussed in the next sections.
%is further strengthened in the next section which shows that the generic models for POLSAR presented in this paper also include the commonly-used models for SAR as its special 1-dimensional case. 
%This result shows how
%In short, these random variables are shown to follow consistent and fixed distributions,
%  regardless of the underlying signal $\Sigma_v$.

@@

%The previous sections have introduced and validated the theoretical models for 3-dimensional ($d=3$) full polarimetric and 2-dimensional ($d=2$) partial polarimetry, cases.
This section shows that the proposed generic model is equally  applicable to the 1-dimensional ($d=1$) case,
  which is physically equivalent to  collapsing the multi-dimensional POLSAR dataset  into conventional single dimensional SAR data.
Mathematically, the sample covariance matrix $C_v$ is reduced to the sample variance while the determinant $|C_v|$ is equivalent to  the scalar variance value.
As variance is equal to intensity $I$ for SAR data, our result is consistent with previous results for SAR intensity data.
Hence this derivation can be considered
  both as cross-validation for the proposed POLSAR model
  as well as evidence showing that
    the proposed generic model for POLSAR also includes the traditional model for SAR intensity in its natural coverage.
%  as well as reminding us that SAR is a special case of POLSAR.

@@

Upon setting $d=1$ into the above equations,
  the equations become:
\begin{align*}
  |C_v| &\sim  \frac{|\Sigma_v|}{(2L)} \cdot \chi (2L) \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim \frac{1}{(2L)} \cdot \chi(2L) \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
\mathbb{R}_{C} = \frac{C_1}{C_2} &\sim \prod_{i=0}^{d-1} \frac{\chi(2L-2i)}{\chi(2L-2i)}
\end{align*}

@@

\section{Unifying the different discrimination measures proposals for bpth SAR and POLSAR}

At least for both fields of SAR and POLSAR, statistical models have always been the foundation to derive discrimination measures.
For SAR, the statistical model for SAR intensity has been used to derive the most widely used intensity-ratio discrimination measure \cite{Rignot_1993_TGRS_896}.
For POLSAR, the same case should apply, 
  except that so far only asymptotic distribution has been derived for the most common foundation, i.e. the likelihood test statistics.
With the insight gained from section \ref{sec:sar_special_case_of_polsar}, 
  it should be clear that the proposed model can also be used to derive discrimination measures for POLSAR.

As a matter of course, in view of the models given in Eqns \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.

As a by-product of this exact derivation,
  several discrimination measures for the common case of $L_x=L_y$ are also proposed.
They are the determinant-ratio and the change-ratio presented in Section \ref{sec:theoretical_model}.
Compared to existing discrimination measures for POLSAR reviewed in section \ref{sec:lit_review}, 
  the proposed dissimilarity measures are much simpler.
They, at the same time, are multi-dimensional extensions of the intensity-ratio discrimination measure widely used in SAR.

Thus the core of the proposed approach is that the determinant of the POLSAR covariance matrix is proposed as the scalar and representative observable of the complex and multi-dimensional data.
%What is the novelty of your solution? What are its advantages
Compared to other published scalar observables with accompanying statistical model,
  the proposed observable and its models are highly representative of the multidimensional data.
The representative power of the proposed observable is justified by the following reasons.
Firstly, the covariance matrix determinant, when collapsed into one-dimensional case, transforms neatly into the representative SAR intensity.
Secondly, the statistical model for this observable is shown as the generic multi-dimensional extension of the traditional and representative intensity model for the one dimensional SAR data.
And last but not least, this observable also leads to the determinant-ratio discrimination measure which
  not only is simpler in concept and computation than the existing POLSAR discrimination measures
  but also can be considered as the natural multi-dimensional extension of the commonly-used intensity-ratio discrimination measure in SAR.
% Prof Vun Comment: already mentioned above, no need to repeat
%hai thoughts: actually this should be included to justify our title, and our approach. It is however being moved to discussion / conclusion section
However, before further discussion ensues, 
  it is necessary to validate the models presented with practical data.

\section{Validating the proposed models against real-life data}
\label{sec:polsar_models_validation}

This section aims to verify the theoretical models given in Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:determinant_ratio_distribution} and \ref{eqn:change_ratio_distribution} against practical data.
All of these models require the estimation of two parameters from the captured data.
These are (i) the dimensional number $d$ and (ii) the look number $L$.
The dimensional number $d$ is related to the type of (POL)SAR data captured,
  with $d=1,2,3$ corresponding to the cases of SAR, partial and full polarimetric SAR data, respectively.
The look-number $L$ can be taken either as nominally stated by the data provider,
  or it can be estimated using the technique proposed by \cite{Anfinsen_2009_TGRS_3795}.

To show the robustness of the proposed models, 
  their validations are also put to test on different types of POLSAR sensors.
Specifically in this paper, the air-borne four-look ($L=4$) AIRSAR image of Flevoland and
  a fine-quad single-look complex RADARSAT2 image ($L=1$) are used.
Since the determinant of the covariance matrix is only significant on multi-look data,
  nine-look processing is first applied on the single-look RADARSAT2 data ($L=9$).

\subsection{The Traditional case of SAR ($d=1$)}

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for a homogeneous area in both the AIRSAR Flevoland dataset is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained with ENL set an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
%In this case, even though the RADARSAT2 has a nominal look-number of 9, its ``true'' value is estimated as $L=$ (TODO:CALC).
In all cases, good visual match is apparent, which hence validates the model proposed above. %providing us a simple validation of the result.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[AIRSAR (HH) determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_determinant_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_determinant}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_determinant_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_determinant}
	} \\
	\subfloat[AIRSAR (HH) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_det_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_det_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	} \\
	\subfloat[AIRSAR (HH) change ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_change_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
	} 
	\hfill	
	\subfloat[RADARSAT2 (HH) change ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	}
\end{tabular}
\caption{The specific d=1 models are validated on both practical RADARSAT2 and AIRSAR datasets}
\label{fig:verify_POLSAR_model_1D}
\end{figure}

\subsection{The Multi-dimensional case of POLSAR ($d=2,3$)}

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar tests %like the one described for SAR data 
  are carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image using the technique described in \cite{Anfinsen_2009_TGRS_3795}. %and is also noted in the plots.
From the plots, their visual matches again validates the proposed model.

@@

%\subsection{Theoretical Implications of This Proposal}
\section{Discussion}
\label{sec:discussion}

%Let us begin by noting
To begin with a few theoretical properties of the proposed statistical models are discussed.
First, the use of covariance matrix log-determinant may be related to the standard eigen-decomposition method of the POLSAR covariance matrices.
In fact, the log-determinant can also be computed as the sum of log-eigenvalues.
Specifically $\ln{|M|} = \sum \ln{\lambda_M}$ where $\lambda_M$ denotes all the eigenvalues of M.
Thus similar to other eigenvalue based approach (e.g. entropy/anisotropy, ...),
  the models presented here are invariant to polarization basis transformations.

@@

Intuitively, the concept of the determinant being representative of the POLSAR data can be thought of as being similar to the concept of using the magnitude as being representative of a complex number
(in a sense this is also why intensity is widely considered representative of the SAR data complex).
Similar to the determinant, the magnitude observable is also scalar and is not lossless.
To fully represent the complex number, another observable is needed besides the magnitude.
The magnitude however is much more widely used, as it can be shown that it is invariant to a change in the reference frame.
Similarly, the determinant of the POLSAR covariance matrix is also scalar and invariant to a change of polarization basis.
Thus while it is not fully representative of the POLSAR data, it can be said to be highly representative of it.
The important thing is that, it can be used to give a scalar comparison of the multi-dimensional data.
This is also similar to the most commonly used magnitude to compare two or more complex numbers. %, most of the times the magnitude is to be used.

Besides the above properties, the theoretical models may also provide an alternative derivation for the likelihood test statistics widely used in POLSAR.
Similar to the way that other measures of distance can be used to derive POLSAR classifiers \cite{Lee_1999_TGRS}, change detectors \cite{Conradsen_2003_TGRS_4}, edge detectors \cite{Schou_2003_TGRS_20} or other clustering and speckle filtering techniques \cite{Le_2010_ACRS} \cite{Le_2011_ACRS}, 
new detection, classification, clustering or speckle filtering algorithms can be derived using the models presented in this paper.
These applications however are not included in this paper, but rather left as a suggestion for future investigations.
There are a few reasons for this decision.
The first is that the advantages of the proposed models have been demonstrated through a few applications in this paper:
  the first being a faster ENL evaluation for POLSAR data and
  the second being the visual evaluation of POLSAR speckle filters.
The second reason is that such an application would be quite big in scope, and as such would easily distract the reader from the main proposal of this paper.
Last but not least, this paper focuses on the proposal of several scalar and representative statistical models for POLSAR, which results in associated statistical models and discrimination measures.
It does not, technically speaking, focuses on proposing new discrimination measures.

@@
